Neighbor-Based Decentralized Training Strategies for Multi-Agent Reinforcement Learning

### 2024-10-02 paper-2024-sac-marl-neighbouring-strategies.tex
chore: on formalization

\begin{abstract}
Multi-agent reinforcement learning  is gaining popularity as a powerful extension of
  single-agent reinforcement learning for training collaborative or competitive agents
  in complex environments. 
%
MARL introduces unique challenges, such as non-stationarity, partial observability, 
  and credit assignment that complicate the learning process.  
%
Typically, in the literature, centralized training methods are used to tackle some of these challenges, 
  although they often face scalability limitations as the number of agents grows.
%
For this reason, decentralized training techniques have been proposed, however, 
  generally they have a longer convergence time compared to centralized ones 
  and may lead to instability.
%
This paper investigates the effectiveness of various neighbor-based decentralized training strategies 
  as a viable alternative to centralized training. 
%  
We evaluate Experience Sharing, Nearest Neighboring Averaging, and Nearest Neighboring Consensus methods 
  in a custom multi-agent environment and compare their performance against centralized training. 
%  
Our results show that neighbor-based methods can achieve comparable performance to centralized training 
  while offering improved scalability and communication efficiency. 
%
% Finally, we discuss the trade-offs between these methods and provide insights into their applicability 
%   in different scenarios.
\end{abstract}

%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

@@

Therefore, 
this paper proposes and formalizes a family of distributed learning methodologies based on neighborhood information. 
This approach aims to: 
\begin{itemize}
  \item incorporate richer information compared to independent learners, and
  \item facilitate the emergence of collective behaviors through local interactions.
\end{itemize}
Specifically, we investigate three distinct neighboring-based training strategy encompassing both neural network-based approaches 
(e.g., neighboring averaging and consensus) 
and experience-based methods inspired by centralized experience sharing. 
We demonstrate the effectiveness of these strategies in comparison to both centralized and fully distributed approaches, 
highlighting how neighborhood-based policies offer a compelling compromise between these two extremes.

@@

\section{Background}\label{sec:background}
\begin{itemize}
\item Introduce Multi-Agent Reinforcement Learning (MARL) and its specific challenges.
\item Provide a formalization (networked agents?) and explain the key concepts: joint action space, shared reward, partial observability, and credit assignment.
\item Explain different MARL training schemes: centralized training (CTCE, CTDE), distributed training (DTDE).
\end{itemize}

>>>>

\begin{abstract}
Multi-agent deep reinforcement learning (MADRL) is gaining popularity as a powerful extension of
  single-agent reinforcement learning for training collaborative or competitive agents
  in complex environments. 
%
MADRL, however, introduces unique challenges, such as non-stationarity, partial observability, 
  and credit assignment that complicate the learning process.
%
Typically, in the literature, centralized training methods are used to tackle some of these challenges, 
  although they often face scalability limitations as the number of agents grows.
%
For this reason, decentralized training techniques have been proposed, however, 
  generally they have a longer convergence time compared to centralized ones 
  and may lead to instability.
%
This paper investigates the effectiveness of various neighbor-based decentralized training strategies based on the well-known deep-q learning algorithm 
  as a viable alternative to centralized training. 
%  
We evaluate experience sharing, nearest neighboring averaging, and nearest neighboring consensus methods 
  in a custom multi-agent environment and compare their performance against centralized training. 
%  
Our results show that neighbor-based methods can achieve comparable performance to centralized training 
  while offering improved scalability and communication efficiency. 
%
% Finally, we discuss the trade-offs between these methods and provide insights into their applicability 
%   in different scenarios.
\end{abstract}

%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_sigagentnificance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

@@

Therefore, 
this paper proposes and formalizes a family of distributed learning methodologies based on neighborhood information---in this research area is typically referred to as \emph{networked agents}.
This approach aims to: 
\begin{itemize}
  \item incorporate richer information compared to independent learners, and
  \item facilitate the emergence of collective behaviors through local interactions.
\end{itemize}
Specifically, we investigate three distinct neighboring-based training strategy encompassing both neural network-based approaches 
(e.g., neighboring averaging and consensus) 
and experience-based methods inspired by centralized experience sharing. 
We demonstrate the effectiveness of these strategies in comparison to both centralized and fully distributed approaches, 
highlighting how neighborhood-based policies offer a compelling compromise between these two extremes.

@@

\section{Background and Motivation}\label{sec:background}
Multi-agent reinforcement learning is an extension of a sequientail decision-making process, 
where multiple agents interact with each other and the environment.
In particular, each agent in the system aims to maximize its own reward,
which is typically a function of the global state (sometime not known, aka partial observable) and the joint action taken by all agents.
%
In doing that, each agent perceive the environment through its own observation function,
which maps the global state to the agent's local observation and perform an action based on its policy.
%
This action is then combined with the actions of other agents to form a joint action,
which is used to update the global state and reward.

MADRL is a subfield of reinforcement learning that focuses on training multiple agents to interact with each other within a shared environment.
%
In this \emph{deep} variant, agents are typically represented by neural networks that learn to maximize a shared (or individual) reward signal.
% 
%
MADRL is a wide area, and it encompassed a variety of scenarios, 
such as cooperative (e.g., multi-agent pathfinding), 
competitive (e.g., multi-agent games), 
and mixed environments.
%
In this paper, we focus on cooperative scenarios, where agents must collaborate to achieve a common goal,
introducing a common formalization (networked agents) and discussing the key concepts of joint action space, shared reward, partial observability, and credit assignment, following by different training schemes (centralized and distributed), with a discussion on the state of the art and motivation for our work.
\subsection{Formalization}
Since MARL is a broad field, it is essential to provide a formalization to guide the discussion.
In this paper, we consider a partially observable networked multi-agent system (PONMAS) as a tuple $(\mathcal{G}, \mathcal{S}, \mathcal{A}, \mathcal{O}, \mathcal{P}, \mathcal{R}, \gamma)$, where:
$\mathcal{G} = (N, E)$ is a communication graph, where $N = {1, \dots, n}$ is the set of $n$ agents and $E \subseteq N \times N$ represents the communication links between agents. Time-varying graphs $\mathcal{G}_t = (N, E_t)$ can be used to represent communication evolving over time $t$.
%
$\mathcal{S}$ is the global state space.
%
$\mathcal{A} = \mathcal{A}^1 \times \dots \times \mathcal{A}^n$ is the joint action space, where $\mathcal{A}^i$ is the action space of agent $i$.
%
$\mathcal{O} = \mathcal{O}^1 \times \dots \times \mathcal{O}^n$ is the joint observation space, where $\mathcal{O}^i$ is the observation space for agent $i$. 
%
Observations $o^i \in \mathcal{O}^i$ provide partial information about the global state.
$\mathcal{P}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \to [0, 1]$ is the state transition function, describing the probability of transitioning to a new state $s' \in \mathcal{S}$ given the current state $s \in \mathcal{S}$ and joint action $a \in \mathcal{A}$.
%
$\mathcal{R} = {\mathcal{R}^i}_{i \in N}$, where $\mathcal{R}^i: \mathcal{S} \times \mathcal{A} \to \mathbb{R}$ is the reward function for agent $i$.
%
$\gamma \in [0, 1]$ is the discount factor.
%
Each agent $i$ at time $t$ receives an observation $o^i_t \in \mathcal{O}^i$, takes an action $a^i_t \in \mathcal{A}^i$ based on its local policy $\pi^i: \mathcal{O}^i \times \mathcal{A}^i \to [0,1]$, and receives a reward $r^i_{t+1} = \mathcal{R}^i(s_t, a_t)$. 
%
The global state evolves according to $\mathcal{P}$. 
Agents can exchange information with their neighbors in $\mathcal{G}$ (or $\mathcal{G}t$ if the graph is time-varying). 
%
The objective is typically to maximize the expected discounted sum of rewards $\sum{t=0}^{\infty} \gamma^t r^i_t$ for each agent $i$, 
or some global reward function like the average reward over all agents.
Formally, it can be defined as:
\[ J(\{\pi^i\}) = \lim_{T \to \infty} \mathbb{E} \left[ \frac{1}{T} \sum_{t=0}^{T-1} \frac{1}{N} \sum_{i \in \mathcal{N}} R^i(s_t, a_t) \right] \]

In doing this, agents typically use other function to estimate the value of taking an action in a given state, such as the action-value function or Q-value. 
The latter represents the expected cumulative discounted reward the agent will receive by taking action $a \in \mathcal{A}$ in state $s \in \mathcal{S}$ and then following its policy $\pi^i$:

\[ Q^i(s, a) = \mathbb{E}_{\pi^i} \left[ \sum_{t=0}^{\infty} \gamma^t r^i_{t+1} | s_0 = s, a_0 = a \right] \]

Similar to the individual reward function, a global average reward function can be defined as $\bar{\mathcal{R}}(s, a) = \frac{1}{n} \sum_{i=1}^n \mathcal{R}^i(s,a)$, leading to a global average Q-value:

\[ \bar{Q}(s, a) = \mathbb{E}_{\pi} \left[ \sum_{t=0}^{\infty} \gamma^t \bar{\mathcal{R}}(s_t, a_t) | s_0 = s, a_0 = a \right] \]

where $\pi = \{\pi^i\}_{i \in N}$ represents the joint policy.  The goal in collaborative MARL is often to find a joint policy that maximizes this global average Q-value or a similar global objective.
%\subsection{Key Concepts}

\subsection{Training Schemes}

\subsection{State of the Art}

\subsection{Motivation}

\begin{itemize}
\item Introduce Multi-Agent Reinforcement Learning (MARL) and its specific challenges.
\item Provide a formalization (networked agents?) and explain the key concepts: joint action space, shared reward, partial observability, and credit assignment.
\item Explain different MARL training schemes: centralized training (CTCE, CTDE), distributed training (DTDE).
\end{itemize}
